---
title: "Replication"
#author: "JP"
#date: "January 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Day 1 - Concepts and Measures

The research question is "Does being an incumbent in Brazil improve your performance at the next election?". The research methodology is a regression discontinuity comparing close winners and losers. 

1. What is the data source and sample used to answer the research question? What alternative data samples might have been used?

2. What is the unit of analysis? What alternative units of analysis could have been used?

3. What is the measure of "performance at the next election" used in the code? What alternative measures could have been used?

4. How is the concept of 'incumbent' measured? Are their any risks with this measure?

5. How is the comcept of 'close winner and loser' measured in the code? What alternative measures could have been used?

What dataset are we testing this argument on?
What are our units of analysis? Parties/candidates/coalitions?
How do we measure the concepts? Winning vote margin in number/pct
What units are the data in?
How do we clean the data?
Missing data?

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(cepespR)

d_2000 <- get_elections(year=2000, position="Mayor", regional_aggregation="Municipality", political_aggregation="Candidate")

d_2004 <- get_elections(year=2004, position="Mayor", regional_aggregation="Municipality", political_aggregation="Candidate")
```

```{r}
d_2000 <- d_2000 %>% filter(NUM_TURNO==1) %>% 
  group_by(COD_MUN_IBGE) %>% 
  mutate(vote_pct_2000=QTDE_VOTOS/sum(QTDE_VOTOS), Electorate=sum(QTDE_VOTOS)) %>% 
  mutate(rank_2000=rank(-QTDE_VOTOS)) %>%
  arrange(COD_MUN_IBGE,rank_2000) %>% 
  mutate(first_rank_vote_pct=max(vote_pct_2000), second_rank_vote_pct=nth(vote_pct_2000,2)) %>%
  mutate(win_margin=ifelse(rank_2000==1,vote_pct_2000-second_rank_vote_pct,vote_pct_2000-first_rank_vote_pct)) %>%
  mutate(Incumbent=ifelse(rank_2000==1,1,0))
```


## Day 2 - Observational regression

What does basic regression look like? 
What happens when we add control variables?
What happens when we change the selection of data?

## Day 3 - Causal regression

What does the RDD regression look like?
Balance tests
Continuity tests

## Day 4 - Generalizability

Robustness checks to alternative specifics, samples (maybe another year?)
How different are those candidates/muns near the threshold?

## Day 5 - Constructive Critiques

Placebo tests
Heterogeneity tests
Mechanisms