---
title: "Replication"
#author: "JP" & "GFA"
#date: "January 27, 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```

## Day 1 - Concepts and Measures

The research question is "Does incumbency in Brazil improve your performance at the next election?". The research methodology is a regression discontinuity comparing close winners and losers. You have been provided with the raw data and need to prepare it for the analysis (but not do the analysis itself). 

Let’s begin analyzing the information provided in the paper:

1. What is the data sample used to answer the research question? What alternative data samples might have been used?

2. What is the measure of how “close” were the winners and losers to the threshold (of a tied election)? What different measures could have been used?

Now, let’s take a look at the dataset:

3. What is the unit of analysis in the dataset? What alternative units of analysis could have been used?

4. How many observations are in the dataset?

5. Compare the variables of interest reported in the paper with the dataset columns. Which variables in the dataset reflect those discussed in the article?

Let’s start exploring the data. What would the answer to our research question be using an ‘observational’ methodology?

6. Using variables from the dataset, conduct a basic regression of incumbency status on electoral performance in 2004. Interpret the results.
Hint: Generate a table with the regression outputs and compare their coefficients and standard errors.

7.Compare your results with the findings in Titiunik 2011. Are the results consistent? How do they differ?


```{r}

#d <- read_csv("Data_Day_2.csv")
#Mun_data <- read_csv("Mun_data.csv")
#d <- d %>% left_join(Mun_data, by="COD_MUN_IBGE")
#d$UF.y <- NULL
#d <- d %>% rename("UF" = 1)
#write.csv(d, "Data_Replication.csv")

setwd("") #complete with your working directory

d <- read_csv("Data_Replication.csv")
d %>% lm(vote_pct_2004~Incumbent_2000, data=.) %>% summary()
```

## Day 2 - Observational regression will controls


Yesterday we ran the observational regression to answer the question: “Does incumbency in Brazil improve your performance at the next election?”. Today we will consider some reasons why our results might not give us a good answer to that question.

Let’s start with why some observations are ‘treated’ and others ‘control’.

1. Briefly describe the treatment assignment process for all of the units in the data.

2. How is the treatment assignment process different for units close to the discontinuity threshold (where winning margin is close to zero)?

One causal critique might be that there is confounding/omitted variable bias: More effective or popular parties are more likely to become incumbents and also to perform better at the next election.

3. To try and adjust for this, add a control for party (equivalently, a fixed effect for party). How does this affect the results?

4. We probably want a wider range of controls to address other sources of omitted variable bias. Add controls for population, IDHM (Human Development Index) and PIB (GDP per capita) to your regression, one at a time. How does this affect your results?

5. What happens if you add all the controls at the same time?

A different causal critique is that there may be selection bias. Perhaps incumbent parties who realized they are unpopular decided not to run again?

6. We do not have these parties in our 2004 dataset, but to estimate how severe this problem might be, calculate what percentage of incumbent parties who ran in 2000 decided NOT to run again in 2004. Is this high or low?

7. We also want to know whether the parties that did not compete again were different from those that did. We don’t have any polling data from just before the 2004 elections to understand if weak parties were more likely to not run, but we can use the results of the 2000 election to try and identify ‘weaker’ and ‘stronger’ parties. Did parties that ran again in 2004 receive a higher share of the vote in 2000 that those that decided not to run again in 2004?



```{r}
d %>% lm(vote_pct_2004~Incumbent_2000 + SIGLA_PARTIDO, data=.) %>% summary()

```

```{r}

d %>% lm(vote_pct_2004~Incumbent_2000 + POP_201, data=.) %>% summary()
d %>% lm(vote_pct_2004~Incumbent_2000 + IDHM_10, data=.) %>% summary()
d %>% lm(vote_pct_2004~Incumbent_2000 + PIB_PER, data=.) %>% summary()

d %>% lm(vote_pct_2004~Incumbent_2000 + POP_201 + IDHM_10 + PIB_PER, data=.) %>% summary()

```


```{r}
d %>% filter(Incumbent_2000==1) %>% 
  mutate(ran_in_2004=ifelse(is.na(vote_pct_2004),"No","Yes")) %>% 
  group_by(ran_in_2004) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(pct_ran_in_2004=n/sum(n))

d %>% filter(Incumbent_2000==1) %>% 
  mutate(ran_in_2004=ifelse(is.na(vote_pct_2004),"No","Yes")) %>% 
  group_by(ran_in_2004) %>% 
  summarize(avg_vote_share_2000=mean(vote_pct_2000)) 

```


## Day 3 - Causal regression

To overcome the causal problems we saw yesterday, Titiunik implements a regression discontinuity.

1. Implement the regression discontinuity using your measure of ‘close elections’ in 2000, the indicator of incumbency status in 2000 and the measure of electoral performance in 2004.

2. Interpret the findings of the regression discontinuity. How do they differ from the observational results in Day 1?

Now let’s go back and check our assumptions for the regression discontinuity design to see if we should trust these new results:

3. One assumption of our regression discontinuity is that comparing incumbents that just won and lost elections in 2000 will produce ‘balance’ on potential omitted variables. There are thousands of variables we could check, but let’s assess balance on the size of the municipality by comparing the number of voters in 2000 within 5% of a tied election.

4. How does the balance close to the threshold compare with the balance of winners and losers in the full dataset?

5. Check for balance within 5% of the threshold on other characteristics of the municipalities between treated and control units using the variables population, IDHM (Human Development Index) and PIB (GDP per capita).

6. Another assumption of regression discontinuity is that parties cannot manipulate their ‘score’ on the running variable - their vote share. We can assess this by checking for continuity in the distribution of the variable measuring distance to the threshold, in our case winning margin in 2000. Test this assumption by implementing a McCrary density test (DCdensity in the ‘rdd’ library). What do the results show?


```{r}
d %>% lm(vote_pct_2004~Incumbent_2000 + win_margin_2000, data=.) %>% summary()
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
d <- d %>% group_by(COD_MUN_IBGE) %>% mutate(Electorate_2000=sum(QTDE_VOTOS_2000))

balance_test_electorate <- d %>% filter(win_margin_2000<0.05 & win_margin_2000>-0.05) %>% 
  group_by(Incumbent_2000) %>% 
  summarise(Electorate_3pct=mean(Electorate_2000))

balance_test_electorate_all <- d %>% 
  group_by(Incumbent_2000) %>% 
  summarise(Electorate_all=mean(Electorate_2000))

balance_test_pop <- d %>% filter(win_margin_2000<0.05 & win_margin_2000>-0.05) %>% 
  group_by(Incumbent_2000) %>% 
  summarise(POP_201_2=mean(POP_201))

balance_test_IDHM <- d %>% filter(win_margin_2000<0.05 & win_margin_2000>-0.05) %>% 
  group_by(Incumbent_2000) %>% 
  summarise(IDHM_10_2=mean(IDHM_10))

balance_test_PIB <- d %>% filter(win_margin_2000<0.05 & win_margin_2000>-0.05) %>% 
  group_by(Incumbent_2000) %>% 
  summarise(PIB_PER_2=mean(PIB_PER))
```

```{r}
install.packages("rdd")
library(rdd)
mccrary <- DCdensity(d$win_margin_2000,cutpoint=0,verbose=TRUE)
```

## Day 4 - Generalizability

How much do we learn from the regression discontinuity results estimated in Day 3? First, we will implement some robustness checks:

1. Our regression discontinuity assumed a *linear* relationship between distance to a tied election and subsequent electoral performance. Implement an alternative regression discontinuity with a more flexible quadratic relationship. How do the results differ?

2. An alternative way to implement the regression discontinuity is to limit the data to only those units 'close' to the threshold of a tied election. Remove any data that is more than +/-5\% from the threshold and run your (linear) regression discontinuity again. How do the results change?

3. Estimating the uncertainty (standard errors) for regression discontinuities is tricky. The ‘correct’ way to do it has been programmed in the ‘rdrobust’ package. Implement the (linear) regression discontinuity in rdrobust and compare the standard errors.

4. If our theory is correct, we should only find an effect at the threshold (distance=0). Use the ‘RDestimate’ function (in the ‘rdd’ library) to perform placebo tests with the threshold set to distance=0.1,-0.1,0.05 and -0.05. What do we learn from this test?

What can we learn about the scope of our conclusions?


5. To assess how different units near the threshold might be, let's compare which parties are most represented near to the threshold (+/-5\%) with those that win or lose by a landslide (win or lose by more than 20\%).

6. Based on the above tests, how generalizable do you believe the findings in Titiunik 2011 to be?

*Extra task*: We can further explore the robustness and generalizability of our findings by replicating the study with different data samples. Use the data for the 2012 and 2016 elections to implement a new (linear) regression discontinuity. Is the relationship the same as the results you estimated for 2000-04? 

```{r}
d %>% lm(vote_pct_2004~Incumbent_2000 + win_margin_2000 + I(win_margin_2000^2), data=.) %>% summary()

```

```{r}
window <- 0.05
d_subset <- d %>% filter(win_margin_2000<window & win_margin_2000>-window)

d_subset %>% lm(vote_pct_2004~win_margin_2000+Incumbent_2000,data=.) %>% summary()
```

```{r}
install.packages("rdrobust")
library(rdrobust)
robust <- rdrobust(d$vote_pct_2004, d$win_margin_2000, c=0, all=T)
robust$coef
robust$se
```

```{r}
RDestimate(vote_pct_2004~win_margin_2000, data=d, cutpoint=0) %>% summary()

RDestimate(vote_pct_2004~win_margin_2000, data=d, cutpoint=0.1) %>% summary()
RDestimate(vote_pct_2004~win_margin_2000, data=d, cutpoint=-0.1) %>% summary()
RDestimate(vote_pct_2004~win_margin_2000, data=d, cutpoint=0.05) %>% summary()
RDestimate(vote_pct_2004~win_margin_2000, data=d, cutpoint=-0.05) %>% summary()
```


```{r}
d %>% filter(win_margin_2000<0.05 & win_margin_2000>-0.05) %>% 
  group_by(SIGLA_PARTIDO) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(Pct=n/sum(n)) %>% 
  arrange(-Pct)

d %>% filter(win_margin_2000>0.2 | win_margin_2000 < -0.2) %>% 
  group_by(SIGLA_PARTIDO) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(Pct=n/sum(n)) %>% 
  arrange(-Pct)

```


```{r}
#d_2012 <- get_elections(year=2012, position="Mayor", regional_aggregation="Municipality", political_aggregation="Party")
#d_2012_processed <- d_2012 %>% select(ANO_ELEICAO,NUM_TURNO,UF,COD_MUN_IBGE,NOME_MUNICIPIO,SIGLA_PARTIDO,QTDE_VOTOS)
#d_2012_processed %>% write_csv("2012 Mayoral Elections Data.csv")
#d_2012 <- read_csv("2012 Mayoral Elections Data.csv")

#d_2016 <- get_elections(year=2016, position="Mayor", regional_aggregation="Municipality", political_aggregation="Party")
#d_2016_processed <- d_2016 %>% select(ANO_ELEICAO,NUM_TURNO,UF,COD_MUN_IBGE,NOME_MUNICIPIO,SIGLA_PARTIDO,QTDE_VOTOS)
#d_2016_processed %>% write_csv("2016 Mayoral Elections Data.csv")
#d_2016 <- read_csv("2016 Mayoral Elections Data.csv")

#d_2012 <- d_2012 %>% filter(NUM_TURNO==1) %>% 
#  group_by(COD_MUN_IBGE) %>% 
#  mutate(vote_pct_2012=QTDE_VOTOS/sum(QTDE_VOTOS)) %>% 
#  mutate(rank_2012=rank(-QTDE_VOTOS)) %>%
#  arrange(COD_MUN_IBGE,rank_2012) %>% 
#  mutate(first_rank_vote_pct=max(vote_pct_2012), second_rank_vote_pct=nth(vote_pct_2012,2)) %>%
#  mutate(win_margin=ifelse(rank_2012==1,vote_pct_2012-second_rank_vote_pct,vote_pct_2012-first_rank_vote_pct)) %>%
#  mutate(Incumbent=ifelse(rank_2012==1,1,0))

#d_2016 <- d_2016 %>% filter(NUM_TURNO==1) %>% 
#  select(COD_MUN_IBGE,SIGLA_PARTIDO,QTDE_VOTOS) %>% 
#  group_by(COD_MUN_IBGE) %>% 
#  mutate(vote_pct_2016=QTDE_VOTOS/sum(QTDE_VOTOS))

#d_alt <- d_2012 %>% left_join(d_2016,by=c("COD_MUN_IBGE","SIGLA_PARTIDO"))
#write.csv(d_alt, "Data Replication 2012 2016.csv")

d_alt <- read_csv("Data Replication 2012 2016.csv")
d_alt %>% lm(vote_pct_2016~Incumbent, data=.) %>% summary()
d_alt %>% lm(vote_pct_2016~Incumbent + win_margin, data=.) %>% summary()
```

```{r}
# Extra: One theory for why incumbents may do better at the next election is due to access to state resouces. So we might expect the incumbency effect to be more positive (less negative) where state resources are more important, i.e. in the North/Northeast, and more negative in the Southeast. Implement a heterogeneity test to compare the results for just Minas Gerais (MG) and Piaui (PI).

install.packages("broom")
library(broom)

d %>% filter(UF %in% c("MG","PI")) %>% 
  group_by(UF) %>%
  do(tidy(lm(vote_pct_2004~win_margin_2000+Incumbent_2000,data=.))) %>%
  filter(term=="Incumbent_2000") %>%
  arrange(p.value)


```

```{r}
# Extra: Preparing a map with winning margins (green = within 3% margin)

install.packages("sf")
library(sf)
install.packages("sp")
library(sp)
mun_shp <- readRDS("mun2.rds") %>% st_as_sf()

d_limited <- d %>% filter(rank_2000==1) %>% mutate(within_3=ifelse(win_margin_2000>-0.03 & win_margin_2000<0.03,1,0))

mun_shp <- mun_shp %>% mutate(COD_MUN_IBGE=as.numeric(as.character(GEOCOD_CH))) %>% left_join(d_limited, by="COD_MUN_IBGE")
mun_shp %>% mutate(within_3=factor(within_3)) %>% ggplot() + geom_sf(aes(fill=within_3), colour=NA) +
  scale_fill_manual(values=c("white","dark green")) +
  theme(legend.position = "none")
ggsave("rd_LATE_map.png", dpi=600)
```

